Cap-and-trade Model for the Chilean Energy Market

### 2023-08-18 docs/Apuntes/cnt_v2.tex

\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage[gen]{eurosym}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[dvipsnames]{xcolor}
\usepackage{graphicx}
\usepackage[stable]{footmisc}
\usepackage{caption}
\usepackage{subcaption} 
\usepackage[title]{appendix}
\usepackage{multirow}
\usepackage{xcolor}
\usepackage[margin=1.0in]{geometry}
\usepackage{array}
%\usepackage{hyperref}

\usepackage{tablefootnote}
\usepackage{lineno}
\usepackage{pifont}% http://ctan.org/pkg/pifont
\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%

@@

\newcommand\boldred[1]{\textcolor{red}{\textbf{#1}}}
\newcommand{\seba}{\textcolor[rgb]{0,1,1}}
\newcommand{\fel}[1]{{\color{red}({\bf FF:}#1)}}
\newcommand{\pia}{\textcolor[rgb]{0,0,1}}

@@

\begin{document}
\linenumbers
\maketitle

\section{Motivation}

Considerar un mercado de cap-and-trade como opción a incrementar el tax en chile\footnote{pag 64. https://mma.gob.cl/wp-content/uploads/2015/09/hojaderuta.pdf}

\section{Model}\label{model}

Let us consider a market of $N$ producers indexed by $i\in\{1,...,N\}$, each of them minimizing the cost of operation and investment. We will assume perfect competition among the producers. Our model will assume two basic stages: operation today and investment for the future in period $t=0$ and the uncertain future where there is a spot market and trade of emission allowances in periods $t\in T:=\{1,...\bar{t}\}$. Uncertainty is revealed in period $t=1$ and is represented by a state of nature $\omega\in\Omega:=\{1,..,K\}$. Thus, for a given state of nature $\omega\in\Omega$, the pair $(t,\omega)$ represents the future in period $t$ when the state $\omega$ is achieved.

@@

At first stage, we consider the present operating plant of the producer $i$, the cost of the operation $C_i$, and the quantity produced $Q_i(t)$ with $t=0$. The selling price (that satisfies the demand) at this stage is $\pi^d(t)$ with $t=0$. The maximum quantity of production at current stage ($t=0$) per producer is $\bar{Q}_i$. Each producer buys a certain amount of carbon allowances $A_i\in\mathbb{R}_+$ at a price $\pi^{a}$ from an auctioneer or regulatory agent.  The auctioneer will try to maximize the revenue obtained from the sold allowances. At this stage when $t=0$, the producer $i$ decides the amount $x_i(t)$ of capacity expansion at a cost $I_i$ for the uncertain future.

At the second stage for each $t\in T$ and given $\omega\in \Omega$, the producer wants to operate in the most efficient way at the spot market choosing production level $Q_i(t,\omega)\in\mathbb{R}_+$ with a cost  $TC_i(t,\omega)\cdot C_i$ where $TC_i(t,\omega)\in\mathbb{R}_+$ is the change of the cost. Also, the producer can invest in a capacity expansion $x_i(t,\omega)\in\mathbb{R}_+$ with an investment cost of $TCR_i(t,\omega)I_i$  where $TCR_i(t,\omega)\cdot C_i$ the change in the investment cost. In addition,  we will consider the existence of a permit trading system, where producers can purchase $P_i(t,\omega)\in\mathbb{R}_+$ permits from other producers if they need to surpass the emission allowances $A_i$, or sell $V_i(t,\omega)\in\mathbb{R}_+$ unused permits if their emissions are below allowances $A_i$. The price of transaction in this system is $\pi^v(t,\omega)$. The producer will evaluate the cost of the uncertain future as a risk neutral agent, i.e., expectated value with a probability obtained from real world data.
\smallskip

The maximum budget for emissions is denoted by  $CAP\in\mathbb{R}_+$ and is drawn from a normal distribution, i.e., $CAP \thicksim N(\mu, \sigma^2)$. This $CAP$ is determined by the auctioneer and  accounts for the total emissions throughout the whole period considered, that is, does not depend on $t$.

\smallskip
The amount of total allowances provided by the auctioneer, $\theta\in\mathbb{R}_+$, is an endogenous variable in our model, and it must satisfy the condition $\sum_i A_i \leq \theta$. 
Since $\theta$ must not surpass the budget for emissions, we consider a small margin of risk given by a probability bound $M\in(0,1)$ for this to happen, i.e.,
\begin{equation}
    Pr(\theta \geq CAP) \leq M,
\end{equation}

and since the variable $CAP$ is chosen from a normal distribution with mean $\mu$ and variance $\sigma^2$, we can write the equation above as
\begin{equation}
    \theta \leq \phi^{-1}(M) \sigma^2 + \mu
\end{equation}

@@

\subsection{Producer's problem}

Each producer $i$ represents one and only one technology in the economy. The choice set of the producer $i$ over a time horizon of $\bar{t}$ years is given by: i) an amount of capacity expansion $x_i:=\left(x_i(0),x_i(t,\omega)_{(t,\omega)\in T\times\Omega}\right)\in\mathbb{R}_+\times\mathbb{R}_+^{T\times\Omega}$, ii) a production plan $Q_i:=\left(Q_{i}(0),(Q_{i}(t,\omega)_{(t,\omega)\in T\times\Omega}\right)\in\mathbb{R}_+\times\mathbb{R}_+^{T\times\Omega}$, iii) allowances bought in the first period $A_i\in\mathbb{R}_+$, iv) allowances bought in the second period $P_i(t,\omega)\in\mathbb{R}_+^{T\times\Omega}$ and v) allowances sold in the second period $V_i(t,\omega)\in\mathbb{R}_+^{T\times\Omega}$.

We define the cost function for $t\in T_0:=\{0\}\cup T$ by

$$C_i(t,\pi^{d}(t),Q_i(t))={\color{red}\Big(a_iQ_i(t)+\frac{b}{2}Q_i(t)^{2}\Big)-\pi^{d}(t) Q_i(t)}.$$

%_{\seba{\left(C_t^{comb}\cdot C^{esp}+CVnC\right)Q_{i,t}+\underbrace{FixedCost_{i,t}}_{\frac{\alpha\cdot I\cdot x}{algo}}}}

%and for the rest of the periods

%$$C_i(t,\pi^{d}(t),Q_i(t,\omega))={\color{red}\Big(a_iQ_i(t)+\frac{b}{2}Q_i(t)^{2}\Big)-\pi_{d}(t) Q_i(t,\omega)}$$

Therefore the objective function of producer $i$ is given by

\begin{align}
\min_{(x_i,Q_i,A_i,P_i,V_i)\in (\mathbb{R}_+^{T_0\times\Omega} \times\mathbb{R}_+^{T_0\times\Omega} \times \mathbb{R}\times\mathbb{R}_+^{T\times\Omega}\times\mathbb{R}_+^{T\times\Omega})} &  C_i(0, \pi^d(0),Q_i(0))+ A_i \pi^{a} + I_i x_i(0) \nonumber \\ &+\sum_{\omega} Pr(\omega)   \Bigg( \sum_{t>0} \frac{1}{(1+R)^t} \Big[ (TC_i(t,\omega)\cdot C_i(t,\pi^d(t),Q_i(t,\omega) )  \nonumber \\
 & + \sum_{t > 0} TCR_i(t,\omega) \cdot I_i\cdot x_i(t,\omega) \Big] + \pi^v(t,\omega)\cdot \left(P_i(t,\omega)-V_i(t,\omega)\right) \Bigg)  \label{prodFO} \\
     \textrm{s.t \ } \nonumber
\end{align}
\begin{align}
    \Big(CF_i \cdot\tau\Big)  \Bigg[\bar{Q_i} + \sum_{t^{\prime}<\bar{t}} x_i(t^\prime,\omega) + x_i(0)+ \bar{Q}_i(t) \Bigg] - Q_i(t,\omega) & \geq 0  & \forall  \quad \omega, t  > 0 & \quad (\alpha_{i,\omega,t}) \\
    \Big(CF_i\cdot\tau \Big)\bar{Q_i}-Q_{i}(0) & \geq 0  &  \quad & \quad (\kappa_i) \\
 A_{i} -V_i(t,\omega) & \geq  0  & \forall  \quad t,\omega & \quad (\beta_{i,\omega})\\
 A_{i} + \sum_{i}P_i(t,\omega) - V_i(t,\omega)-\sum_{t>0}Q_i(t, \omega)\cdot \varepsilon_{i}-Q_i(0)\varepsilon_{i} & \geq  0  &\forall \quad t,\omega & \quad (\gamma_{i,\omega})\\
 Q_i(0) & \geq  0 & \quad & \quad (\lambda_i)\\
 Q_i(t, \omega & \geq  0   & \forall  \quad \omega, t >0 & \quad (\delta_{i,\omega})\\
  x_i(0) & \geq  0 & \quad & \quad (\xi_i)\\
  x_i(t, \omega) & \geq  0   & \forall  \quad \omega, t >0 & \quad (\varphi_{i,\omega,t})\\
{\color{blue} RP_i - \bar{Q}_i - \bar{Q}_i(t) - x_i(0) - \sum_{t > 0} x_i(t,\omega) } & {\color{blue} \geq 0} & {\color{blue} \forall \quad i,\omega }&  {\color{blue} \quad (\psi_{i,\omega}) }
  \end{align}\\

\subsection{Equilibrium}

\subsection{Computation}

@@

\begin{center}
\begin{tabular}{  l l l l  } 
 \hline
 \textbf{Parameter} & \boldred{Suggested} & \textbf{Units} & \textbf{Explanation} \\ 
 \hline
 \hline
 $I_i$ & & USD/MW & Expansion cost per technology $i$ \\  
 $\bar{Q}_i$ & \boldred{$\bar{X}_i$} & MW & Maximum operation capacity per technology $i$  \\ 
 $C_i$ &  & USD/MWh & Operation cost per technology $i$ \\
 $D^{1}$ & & MWh & Total demand in first stage \\
 $D^{t}_{\omega}$ & & MWh & Total demand in period $t > 1$ \\
 $\varepsilon_i$ & & tCO$_{2}$/MWh  & Emission factor of technology $i$\\
 Prob$_\omega$ & \boldred{$\Pi_\omega$} & & Physical probability of scenario $\omega$ \\
 $\mu$ & & tCO$_{2}$ & Mean of the normal distribution of the CAP \\
  $\sigma$ & & tCO$_{2}$ & Standard deviation of the normal distribution of the CAP \\
  $\epsilon$ & &  & Margin for total emission allowances \\
  $R$ & \boldred{df} & & Discount rate \\
  $TC_i$ & & & Change in the cost of operation per technology \\
  $TCR_i$ & & & Change in the investment cost per technology \\
  $CF_i$ & & & Capacity factor per technology\\
  $\tau$ & & hours&  Number of hours in a year, $\tau=8760$ hours\\
  {\color{blue} $RP_i$ } & & {\color{blue} MW} & {\color{blue} Resource potential per technology }\\
\hline
\end{tabular}
\end{center}

@@

\small
\begin{align}
\min_{x,Q,A,P,V} & \underbrace{{\color{red}\Big(a_iQ_{1,i}+\frac{b}{2}Q_{1,i}^{2}\Big)-\pi^{d,1} Q_{1,i}}}_{\seba{\left(C_t^{comb}\cdot C^{esp}+CVnC\right)Q_{i,t}+\underbrace{FixedCost_{i,t}}_{\frac{\alpha\cdot I\cdot x}{algo}}}} + A_i \pi^{a} + I_i x_i^{1} + \sum_{\Omega} prob_\omega   \Bigg( \sum_{t>1} \frac{1}{(1+R)^t} \Big[ (TC_{i,\tau}\cdot {\color{red}(a_i Q_{i,t,\omega}+\frac{b}{2}Q_{i,t,\omega}^{2})}\nonumber \\
    & {\color{red}-\pi^{d,2}_{\omega} Q_{i,t,\omega}} + \sum_{t > 1} TCR_i \cdot I_ix^{t}_{i,w} \Big] - V_{i,\omega}\pi_{i,w}^v+\sum_{j\neq i}  P_{i,j,\omega}\pi_{j,w}^v\Bigg)  \\
     \textrm{s.t \ } \nonumber
\end{align}
\begin{align}
    \Big(CF_i \cdot\tau\Big)  \Bigg[\bar{Q_i} + \sum_{t^{\prime}<t_{y_i}} x^{t^{\prime}}_{i,w} + x_i^{1}+ \bar{Q}_{i,t} \Bigg] - Q_{i,\omega}^{t} & \geq 0  & \forall  \quad \omega, t  > 1 & \quad (\alpha_{i,\omega,t}) \label{const.1}\\
    \Big(CF_i\cdot\tau \Big)\bar{Q_i}-Q_{i}^{t} & \geq 0  &  \quad t  =1 & \quad (\kappa_i) \label{const.2} \\
 A_{i} -V_{i,\omega} & \geq  0  & \forall  \quad \omega & \quad (\beta_{i,\omega})\\
 A_{i} + \sum_{j \neq i}P_{i,j,\omega} - V_{i,\omega}-\sum_{t>1}Q_{i, \omega}^{t}\varepsilon_{i}-Q_i^{1}\varepsilon_{i} & \geq  0  &\forall \quad \omega & \quad (\gamma_{i,\omega})\\
 Q_i^{1} & \geq  0 & \quad & \quad (\lambda_i)\\
 Q_{i, \omega}^{t} & \geq  0   & \forall  \quad \omega, t >1 & \quad (\delta_{i,\omega})\\
  x_i^{1} & \geq  0 & \quad & \quad (\xi_i)\\
  x_{i, \omega}^{t} & \geq  0   & \forall  \quad \omega, t >1 & \quad (\varphi_{i,\omega,t})\\
{\color{blue} RP_i - \bar{Q}_i - \bar{Q}_{i,t} - x^{1}_i - \sum_{t > 1} x^{t}_{i,w} } & {\color{blue} \geq 0} & {\color{blue} \forall \quad i,\omega }&  {\color{blue} \quad (\psi_{i,\omega}) }
  \end{align}\\
\normalsize

@@

\section{Model comparison with other projections}
Following Chile's Energy Policies 2050 (Hoja de Ruta 2050\footnote{http://www.energia.gob.cl/sites/default/files/hoja\_de\_ruta\_cc\_e2050.pdf}), we can compare the results of our model without considering a cap-and-trade scheme, that is, \textit{bussiness as usual}. This comparison is useful to calibrate the parameters of our model with the results of present years and the projections for future years. Table \ref{table:comparison} shows the comparison between goverment projections and our model for the year 2050.

\begin{table}[h!]
\centering
\begin{tabular}{ |c|c|c|c| } 
\hline
 & & Chile's Energy Policies projections (base-case)& Our model \\
\hline
\multirow{4}{8em}{Energy Production (\%)} & NCRE\tablefootnote{According to Chile's Energy Policies document, NCRE are Solar Photovoltaic and Concentrated Solar Power (CSP), Geothermal and Wind } & 22\%-67\% & 25\% \\ 
 & Hidro generation & 24\%-32\% & 21\% \\ 
& Thermal generation & 9\%-46\% & 46\% \\
\hline
%\multicolumn{3}{l}{\footnotetext{According to Chile's Energy Policies document, NCRE are Solar Photovoltaic and Concentrated Solar Power (CSP), Geothermal and Wind }} \\
\end{tabular}
\caption{Energy Generation from our model compared with Chile's Energy Policies projections for the year 2050}
\label{table:comparison}
\end{table}

\fel{ PIA: Veo en los numeros que estamos siempre en el "peor" caso (menos NCRE y hydro y mas thermal)... en la comparacion, el 22\%. 24\% y 46\% de la Hoja de Ruta corresponden al mismo escenario?}

In addition, other studies have predicted the level of energy generation and emissions. Figures \ref{fig:modelbau} and \ref{fig:gener} shows our model prediction of energy generation per technology in the range 2018-2050 and the projection from the study of Generadoras de Chile (2017)\footnote{http://generadoras.cl/documentos/estudios/actualizacion-de-la-proyeccion-de-emisiones-2017-2030-y-analisis-medidas-de-mitigacion-de-co2-equivalente}

\begin{figure}[ht!]
\begin{center}
\includegraphics[width=\textwidth]{Apuntes/Figures/produccion_per_year_sin_cnt.png}
\caption{Energy production projected per technology by our model for the base year (2018) and the year range 2019-2050} \label{fig:modelbau}   
\end{center}
\end{figure}

\begin{figure}[ht!]
\begin{center}
\includegraphics[width=13cm]{Apuntes/Figures/generadoras.png}
\caption{Projected energy production per technology made by Generadoras de Chile}
\label{fig:gener}
\end{center}
\end{figure}

>>>>

\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage[gen]{eurosym}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[dvipsnames]{xcolor}
\usepackage{graphicx}
\usepackage[stable]{footmisc}
\usepackage{caption}
\usepackage{subcaption} 
\usepackage[title]{appendix}
\usepackage{multirow}
\usepackage{xcolor}
\usepackage[margin=1.0in]{geometry}
%\usepackage{hyperref}

@@

\newcommand\boldred[1]{\textcolor{red}{\textbf{#1}}}
\newcommand{\seba}{\textcolor[rgb]{0,1,1}}

@@

\begin{document}
\maketitle

\section{Simple Model}\label{model}

Let us consider a market of $i$ producers, each of them minimizing the cost of operation and investment. We will assume perfect competition among the producers. Our model will assume two basic stages: operation at present (with investment for the future) and the future, at the spot market and trading of emission allowances.

@@

At first stage, we consider the current operating plant and the cost of the operation $C_i$, and the quantity produce $Q_i^{\tau}$ with $\tau=1$. The price of sale (that satisfies the demand) at this stage is $\pi^{d,\tau}_i$ with $\tau=1$. The maximum quantity of production at current stage (initial) per producer is $\bar{Q}_i$. Each producer buys a certain amount of carbon $A_i$ at a price $\pi^{a}$ from an auctioner of regulatory agent.  The auctioneer will try to maximize the revenue obtained from the sold allowances. At this stage, the producer $i$ decides the amount $x_i$ of capacity expansion at a cost $I_i$ for the uncertain future.
\smallskip

At second stage, in the uncertain future, the producer wants to operate in the most efficient way at the spot market. In addition,  we will consider the existence of a permit trading market, where producers can purchase permits from other producers if they need to surpass the emission allowances, or sell the unused permits if their emissions are below allowances. The producer will evaluate the cost of the uncertain future as risk neutral agents, i.e., as an expectation value and we will assume that the probability is a physical probability, which may be obtained from real world data.
\smallskip

The maximum budget for emissions is the  parameter CAP, which in turn, follows a normal distribution, i.e., $CAP \thicksim N(\mu, \sigma^2)$. This CAP is determined by the auctioneer and  accounts for the total emissions throughout the whole period considered, that is, does not depend of $\tau$.

\smallskip
The amount of total allowances provided by the auctioneer $\theta$ is an endogenous random variable in our model, and it must satisfied the condition $\sum_i A_i = \theta$. 
Since $\theta$ must not surpass the budget for emissions, we consider a low margin (or risk) $R$ for this to happen, i.e.,
\begin{equation}
    P (\theta \geq CAP) \leq R
\end{equation}

and since the variable CAP follows a normal distribution with mean $\mu$ and variance $\sigma^2$, we can write the equation above as
\begin{equation}
    \theta \leq \phi^{-1}(R) \sigma^2 + \mu
\end{equation}

@@

\begin{flushleft}
\textbf{Sets}
\begin{itemize}
    \item[] $i$: number of producers
    \item[] $\omega$: possible scenarios
    \item[] $\tau$: model periods

\textbf{Parameters}
\begin{itemize}
   \item[] $\bar{Q}_i$: maximum operation capacity
   \item[] $I_i$: expansion cost
   \item[] $C_i$: operation cost
   \item[] $D$: total demand (exogenous)
   \item[] $CAP$: maximum budget of emissions 
   \item[] $R$: margin for total emission allowances
 \item[] $\varepsilon_{i}$: Emissions factor of producer $i$
\end{itemize}
\hspace{0.5cm}

\textbf{Variables}
\begin{itemize}
    \item[] $Q^{\tau}_i$: produced quantity
    \item[] $A_i$: emission allowances
    \item[] $P^{\tau}_i$: purchased permits
    \item[] $V^{\tau}_i$: sold permits
    \item[] $\pi^{a}_i$: dual to CAP constraint
    \item[] $\pi^{v}_i$: dual to trading equilibrium
    \item[] $\pi^{d}_i$: dual to fulfilment of the demand
    \item[] $x_i$: capacity expansion decision
    \item[] $\theta$: emission permits available in the market
    \end{itemize}

\textbf{Equations}
\smallskip
The model below is defined for each $i \in \{ 1,...,n\}$. Each producer minimizes cost as follows

\begin{align} \label{eq:simplemodel}
    \min_{x,Q,A,P,V} & C_iQ_i^{1} + A_i \pi^{a} + I_i x_i \nonumber\\
    & + \sum_{\Omega} prob_\omega \Big[ \sum_{\tau>1} (C_i-\pi_{\omega}) Q_{i,\omega}^{\tau}- V_{i,\omega}\pi_{i,w}^v+\sum_{j\neq i} P_{i,j,\omega}\pi_{j,w}^v\Big]\\
     \textrm{s.t \ } \nonumber
\end{align}
\begin{align}
    \bar{Q_i} + x_{i} - Q_{i,\omega}^{\tau} & \geq 0  & \forall & \quad \omega, \tau  > 1 & \quad (\alpha_{i,\omega,\tau}) \\
    \bar{Q_i}-Q_{i}^{\tau} & \geq 0  &  & \quad \tau  =1 & \quad (\kappa_i) \\
 A_{i} -V_{i,\omega} & \geq  0  & \forall & \quad \omega & \quad (\beta_{i,\omega})\\
 A_{i} + \sum_{j \neq i}P_{i,j,\omega} - V_{i,\omega}-\sum_{\tau>1}Q_{i, \omega}^{\tau}\varepsilon_{i}-Q_i^{1}\varepsilon_{i} & \geq  0  &\forall & \quad \omega & \quad (\gamma_{i,\omega})\\
 Q_{i, \omega}^{\tau} & \geq  0   & \forall & \quad \omega, \tau & \quad (\delta_{i,\omega})
\end{align}

\smallskip
The auctioneer maximizes its revenue from the allowances market; we write the problem as a minimization program as follows
\begin{align}
    \min_{\theta} & -\theta \pi^{a} \\
    \textrm{s.t \ } & \theta \leq \phi^{-1}(R) \sigma^2 + \mu \\
    &\theta \geq 0
\end{align}

Finally, we have three market clearing constraints equations for this market

\begin{align}
\textrm{(available allowances)}: &  \ \   \sum_{\tau} \sum_{i} A^{\tau}_{i,\omega} = \theta   & \forall \ \omega & \ \  (\pi^{a})\\
\textrm{(equilibrium in trading market)}: &   \ \  \sum_{i} P_{i,\omega} = \sum_{i} V_{i,\omega} & \forall \ \omega & \ \ (\pi^{v}) \\
\textrm{(fulfillment of the demand)}:  &   \ \  \sum_{i} Q^{\tau}_{i,\omega} = D^{\tau}_{\omega}, & \forall \ \omega, \tau & \ \ (\pi^{d})
\end{align}

\end{flushleft}

\subsection{KKT conditions}

The Lagrangian function for the producer problem for each $i \in \{ 1,...,n\}$ is defined as follows:

\begin{multline}
    \mathcal{L}_i(x,Q,A,P,V) = C_i Q_i^{1} + A_i\pi^{a} + I_i x_i + \sum_{\Omega} prob_\omega \Big[ \sum_{\tau>1} (C_i-\pi_{\omega}) Q_{i,\omega}^{\tau} - V_{i,\omega}\pi_{i,w}^v+\sum_{j\neq i} \nonumber P_{i,j,\omega}\pi_{j,w}^v\Big]\\ + \kappa_{i}\Big[Q_{i}^{1} - \bar{Q_i}\Big] +\sum_{\omega,\tau>1} \alpha_{i,\omega,\tau}\Big[Q_{i,\omega}^{\tau} - (\bar{Q_i}+ x_i)\Big] + \sum_{\omega}\beta_{i,\omega}\Big[V_{i,\omega}-A_i\Big] \\
    + \sum_{\omega}\gamma_{i,\omega} \Big[-A_{i} - \sum_{j \neq i}P_{i,j,\omega} + V_{i,\omega}+\sum_{\tau>1}Q_{i, \omega}^{\tau}\varepsilon_{i}+Q_i^{1}\varepsilon_{i}\Big] - \sum_{\omega, \tau>1}\delta_{i,\omega,\tau} Q_{i,\omega}^{\tau} - \lambda_{i}\Big[Q^{1}_{i}\Big]
\end{multline}

where $\alpha_{i,\omega,\tau}$, $\kappa_i$, $\beta_{i,\omega}$, $\gamma_{i,\omega}$ and $\delta_{i,\omega}$ and $\lambda_i$ are the lagrange multipliers of the constraints.

With the Lagrangian set up, we can now find de KKT conditions for each of the producers $i \in \{ 1,...,n\}$ by making $\nabla \mathcal{L}_{i}=0$, we obtain:

\begin{align}
    I_i - \sum_{\omega,\tau>1} \alpha_{i,\omega,\tau} & = 0 & \forall i  &  \qquad (x_i)\\
    C_{i} + \kappa_i - \lambda_{i} + \sum_{\omega} \gamma_{i,\omega} & = 0 & \forall i  &  \qquad (Q^{1}_{i})\\
    prob_\omega (C_i-\pi_{\omega}) + \alpha_{i,\omega,\tau} + \gamma_{i,\omega} \varepsilon_{i}-\delta_{i,\omega,\tau} & =0 & \forall i, \omega, \tau>1 &  \qquad (Q_{i,\omega}^{\tau})\\
    \lambda - \sum_{\omega}\beta_{i,\omega} - \sum_{\omega}\gamma_{i,\omega} & =0 & \forall i & \qquad (A_{i}) \\
    -prob_\omega \pi^v_{i,\omega} + \beta_{i,\omega}  + \gamma_{i,\omega} & =0 & \forall i, \omega & \qquad (V_{i,\omega}) \\
    prob_\omega \pi^v_{j,\omega} -\gamma_{i,\omega} & = 0 & \forall i, \omega & \qquad (P_{i,j,\omega})
\end{align}

\begin{align}
    -Q_{i,\omega}^{\tau} + \bar{Q_i} + x_i & \geq 0 & \qquad (\alpha_{i,\omega,\tau}) \\
    -\bar{Q_i} + Q^{\tau}_i & \geq 0 & \qquad (\kappa_i)\\
     A_{i} - V_{i,\omega} & \geq 0 & \qquad (\beta_{i,\omega}) \\
     A_{i} + P_{i,\omega} - V_{i,\omega} - \sum_{\tau>1}Q_{i, \omega}^{\tau}\varepsilon_{i} & \geq 0 & \qquad (\gamma_{i,\omega}) \\
    Q_{i,\omega}^{\tau} & \geq 0 & \qquad (\delta_{i,\omega})
\end{align}

\begin{flushleft}
Complementary slackness\\

\textbf{FALTAN los For all i,w,t...etc}
\end{flushleft}

\begin{align}
    (-Q_{i,\omega}^{\tau} + \bar{Q_i} + x_i)\alpha_{i,\omega,\tau} & = 0 \\
    ( -\bar{Q_i} + Q^{\tau}_i)\kappa_i & = 0 \\
    (V_{i,\omega} - A_{i})\beta_{i,\omega} & = 0 \\
    ( A_{i} + P_{i,\omega} - V_{i,\omega} - \sum_{\tau>1}Q_{i, \omega}^{\tau}\varepsilon_{i})\gamma_{i,\omega} & = 0 \\
    Q_{i,\omega}^{\tau}\delta_{i,\omega} & = 0
\end{align}

\begin{align}
    \alpha_{i,\omega,\tau} & \geq 0 \\
    \kappa_i & \geq 0 \\
    \beta_{i,\omega} & \geq 0 \\
    \gamma_{i,\omega} & \geq 0 \\
    \delta_{i,\omega} & \geq 0 
\end{align}

\begin{equation}
    \mathcal{L}(\theta)= -\theta \pi^{a} - \eta (\phi^{-1}(R) \sigma^2 + \mu - \theta) - mu\theta
\end{equation}

where $\eta$ is the lagrange multiplier. Thus, we can obtain the KKT conditions for the auctioneer problem.

\begin{align}
    -\pi^{a} +\eta -\mu & = 0 \\
    \theta & \geq  0 \\
    (\phi^{-1}(R) \sigma^2 + \mu - \theta) \eta & =  0\\
    \eta & \geq  0
\end{align}

This model has been implemented in GAMS. We have considered a two stage problem, where first stage consists in the year range $2017-2018$ and second stage consists in the range $2019-2037$. We have considered each type of energy generation as a producer. The types of technology considered, according to data from the Comisi\'on Nacional de Energ\'ia are: Biogas, Biomasa, Carbon, Cogenercion, Eolica, FuelOil, Gas, Geotermia, Hydro, PetroleoDiesel, Solar. Also, data for Capacity, Cost of investment, demand have been obtained from the same database.

\section{A slightly more complicated model}

We make things a little more general by considering more detailed aspects:

\begin{itemize}
    \item We consider that at first stage, the plant is operating and fulfiling the demand at first stage, at a price $\pi^{d,1}$
    \item Producers can invest in capacity expansion not only at first stage, but also at any point during second stage. 
    \item Producers can change their plant's technology to maintain the production
\end{itemize}

We modify Equation \ref{eq:simplemodel} taken into account the considerations described above; thus, the problem takes the following form:

\begin{align}
    \min_{x,Q,A,P,V} & (C_i-\pi^{1}) Q_i^{1} + A_i \pi^{a} + I_i x_i^{1} + \sum_{\Omega} prob_\omega \Big[ \sum_{\tau>1} (TC_{i,\tau}\cdot C_i-\pi_{\omega}) Q_{i,\omega}^{\tau}- \nonumber \\ 
    &V_{i,\omega}\pi_{i,w}^v+\sum_{j\neq i}  P_{i,j,\omega}\pi_{j,w}^v+ \sum_{\tau > 1} I_ix_{i,\tau,w}\Big]\\
     \textrm{s.t \ } \nonumber
\end{align}
\begin{align}
    \bar{Q_i} + \sum_{\tau^{t}\leq\tau} x_{i,\tau^{t},w} - Q_{i,\omega}^{\tau} & \geq 0  & \forall  \quad \omega, \tau  > 1 & \quad (\alpha_{i,\omega,\tau})\\
    \bar{Q_i}-Q_{i}^{\tau} & \geq 0  &  \quad \tau  =1 & \quad (\kappa_i) \\
 A_{i} -V_{i,\omega} & \geq  0  & \forall  \quad \omega & \quad (\beta_{i,\omega})\\
 A_{i} + \sum_{j \neq i}P_{i,j,\omega} - V_{i,\omega}-\sum_{\tau>1}Q_{i, \omega}^{\tau}\varepsilon_{i}-Q_i^{1}\varepsilon_{i} & \geq  0  &\forall \quad \omega & \quad (\gamma_{i,\omega})\\
 Q_{i, \omega}^{\tau} & \geq  0   & \forall  \quad \omega, \tau & \quad (\delta_{i,\omega})
\end{align}

\section{Model -- Version 6}

@@

\begin{center}
\begin{tabular}{  l l l l  } 
 \hline
 \textbf{Parameter} & \boldred{Suggested} & \textbf{Units} & \textbf{Explanation} \\ 
 \hline
 \hline
 $I_i$ & & USD/MW & Expansion cost per technology $i$ \\  
 $\bar{Q}_i$ & \boldred{$\bar{Y}_i$} & MW & Maximum operation capacity per technology $i$  \\ 
 $C_i$ &  & USD/MWh & Operation cost per technology $i$ \\
 $D^{1}$ & & MWh & Total demand in first stage \\
 $D^{t}_{\omega}$ & & MWh & Total demand in period $t > 1$ \\
 $\varepsilon_i$ & & tCO$_{2}$/MWh  & Emission factor of technology $i$\\
 Prob$_\omega$ & \boldred{$\Pi_\omega$} & & Physical probability of scenario $\omega$ \\
 $\mu$ & & tCO$_{2}$ & Mean of the normal distribution of the CAP \\
  $\sigma$ & & tCO$_{2}$ & Standard deviation of the normal distribution of the CAP \\
  $\epsilon$ & &  & Margin for total emission allowances \\
  $R$ & \boldred{df} & & Discount rate \\
  $TC_i$ & & & Change in the cost of operation per technology \\
  $TCR_i$ & & & Change in the investment cost per technology \\
  $CF_i$ & & & Capacity factor per technology\\
  $\tau$ & & hours&  Number of hours in a year, $\tau=8760$ hours\\
  {\color{blue} $RP_i$ } & & {\color{blue} MW} & {\color{blue} Resource potential per technology }\\
\hline
\end{tabular}
\end{center}

@@

\begin{align}
\min_{x,Q,A,P,V} & \underbrace{{\color{red}\Big(a_iQ_{1,i}+\frac{b}{2}Q_{1,i}^{2}\Big)-\pi^{d,1} Q_{1,i}}}_{\seba{\left(C_t^{comb}\cdot C^{esp}+CVnC\right)Q_{i,t}+\underbrace{FixedCost_{i,t}}_{\frac{\alpha\cdot I\cdot x}{algo}}}} + A_i \pi^{a} + I_i x_i^{1} + \sum_{\Omega} prob_\omega   \Bigg( \sum_{t>1} \frac{1}{(1+R)^t} \Big[ (TC_{i,\tau}\cdot {\color{red}(a_i Q_{i,t,\omega}+\frac{b}{2}Q_{i,t,\omega}^{2})}\nonumber \\
    & {\color{red}-\pi^{d,2}_{\omega} Q_{i,t,\omega}} + \sum_{t > 1} TCR_i \cdot I_ix^{t}_{i,w} \Big] - V_{i,\omega}\pi_{i,w}^v+\sum_{j\neq i}  P_{i,j,\omega}\pi_{j,w}^v\Bigg)  \label{prodFO} \\
     \textrm{s.t \ } \nonumber
\end{align}
\begin{align}
    \Big(CF_i \cdot\tau\Big)  \Bigg[\bar{Q_i} + \sum_{t^{\prime}<t_{y_i}} x^{t^{\prime}}_{i,w} + x_i^{1}+ \bar{Q}_{i,t} \Bigg] - Q_{i,\omega}^{t} & \geq 0  & \forall  \quad \omega, t  > 1 & \quad (\alpha_{i,\omega,t}) \label{const.1}\\
    \Big(CF_i\cdot\tau \Big)\bar{Q_i}-Q_{i}^{t} & \geq 0  &  \quad t  =1 & \quad (\kappa_i) \label{const.2} \\
 A_{i} -V_{i,\omega} & \geq  0  & \forall  \quad \omega & \quad (\beta_{i,\omega})\\
 A_{i} + \sum_{j \neq i}P_{i,j,\omega} - V_{i,\omega}-\sum_{t>1}Q_{i, \omega}^{t}\varepsilon_{i}-Q_i^{1}\varepsilon_{i} & \geq  0  &\forall \quad \omega & \quad (\gamma_{i,\omega})\\
 Q_i^{1} & \geq  0 & \quad & \quad (\lambda_i)\\
 Q_{i, \omega}^{t} & \geq  0   & \forall  \quad \omega, t >1 & \quad (\delta_{i,\omega})\\
  x_i^{1} & \geq  0 & \quad & \quad (\xi_i)\\
  x_{i, \omega}^{t} & \geq  0   & \forall  \quad \omega, t >1 & \quad (\varphi_{i,\omega,t})\\
{\color{blue} RP_i - \bar{Q}_i - \bar{Q}_{i,t} - x^{1}_i - \sum_{t > 1} x^{t}_{i,w} } & {\color{blue} \geq 0} & {\color{blue} \forall \quad i,\omega }&  {\color{blue} \quad (\psi_{i,\omega}) }
  \end{align}\\

@@
